---
title: "Theme Template for ggplot"
author: "Sam Voisin"
date: "December 15, 2018"
output: html_document
---

## Theme Template for Plotting with ggplot

```{r template, echo = TRUE}

library(tidyverse)
library(gridExtra)

### Master Plot Format ###

master_plot_theme <- theme_light() +
  theme(
    # large plot title
    plot.title   = element_text(color  = "black",
                                face   = "bold",
                                size   = 15,
                                family = NULL,
                                hjust  = NULL,
                                vjust  = NULL),
    # x axis label
    axis.title.x = element_text(color  = "black",
                                face   = "bold",
                                size   = 12,
                                family = NULL,
                                hjust  = NULL,
                                vjust  = NULL),
    # y axis label
    axis.title.y = element_text(color  = "black",
                                face   = "bold",
                                size   = 12,
                                family = NULL,
                                angle  = NULL,
                                hjust  = NULL,
                                vjust  = NULL),
    # x axis tick labels
    axis.text.x  = element_text(face   = "bold",
                                color  = "black",
                                size   = 10,
                                family = NULL,
                                angle  = NULL,
                                hjust  = NULL,
                                vjust  = NULL),
    # y axis tick labels
    axis.text.y  = element_text(color  = "black",
                                face   = "bold",
                                size   = 10,
                                family = NULL,
                                angle  = NULL,
                                hjust  = NULL,
                                vjust  = NULL),
    # text for facet strips
    strip.text   = element_text(color  = "black",
                                face   = "bold",
                                size   = 15,
                                family = NULL,
                                hjust  = NULL,
                                vjust  = NULL),
    # background for facet strips
    strip.background = element_rect(color    = "black",
                                    fill     = "white",
                                    linetype = "solid"),
    # major and minor gridlines; can be added back w/ element_line(...)
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position    = "none")

```



```{r example, echo = TRUE}

# simulate some data
simFrame <- data_frame(
  S = rnorm(100),
  A = rpois(100, lambda = 10),
  M = rep(c("A", "B"), 50)
) %>% mutate(M = factor(M))

# without using the template
withoutFmtScatter <- ggplot(data = simFrame,
                            aes(x = S,
                                y = A,
                                color = M))  +
  geom_point(size = 2)  +
  geom_jitter()         + 
  ggtitle("Test Title") +
  labs(x = "Normally Distributed",
       y = "Poisson Distributed")

# using the template
withFmtScatter <- ggplot(data = simFrame,
                         aes(x = S,
                             y = A,
                             color = M))  +
  geom_point(size = 2)  +
  geom_jitter()         + 
  ggtitle("Test Title") +
  labs(x = "Normally Distributed",
       y = "Poisson Distributed") +
  master_plot_theme

# without using the template - faceted
withoutFmtFacet <- ggplot(data = simFrame,
                          aes(x = S,
                              y = A,
                              color = M))  +
  geom_point(size = 2)  +
  geom_jitter()         + 
  facet_wrap(M~.)       +
  ggtitle("Test Title") +
  labs(x = "Normally Distributed",
       y = "Poisson Distributed")

# using the template - faceted
withFmtFacet <- ggplot(data = simFrame,
                       aes(x = S,
                           y = A,
                           color = M))  +
  geom_point(size = 2)  +
  geom_jitter()         + 
  facet_wrap(M~.)       +
  ggtitle("Test Title") +
  labs(x = "Normally Distributed",
       y = "Poisson Distributed") +
  master_plot_theme



```

#### Here we can see the effect of the template on a scatter plot. The base ggplot is on the left; the formatted version is on the right.

```{r scatterExample, echo = TRUE}

grid.arrange(withoutFmtScatter,
             withFmtScatter,
             nrow = 1,
             ncol = 2)

```

#### This is the same template applied to a faceted scatter plot.

```{r facetExample, echo = TRUE}

grid.arrange(withoutFmtFacet,
             withFmtFacet,
             nrow = 1,
             ncol = 2)

```

#### Make theme changes in the same way you would normally. Additions overwrite the template themes. Below is the previous scatter plot with an additional theme object added to change the color of the plot title and axis titles. Notice that the plot title changed color, but the axis labels did not. This is because the master template specifies the colors of each title explicitly. See the final graph for an example.

```{r Modify, echo = TRUE}

withFmtScatter +
  theme(plot.title = element_text(color = "red"),
        axis.title = element_text(color  = "blue"))

withFmtScatter +
  theme(plot.title = element_text(color = "red"),
        axis.title.x = element_text(color  = "blue"),
        axis.title.y = element_text(color  = "blue"))

```


















